### 2018-05-10 cover-letter/cover-letter.tex
adds cover letter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Plain Cover Letter
% LaTeX Template
%
% This template has been downloaded from:
% http://www.latextemplates.com
%
% Original author:
% Rensselaer Polytechnic Institute (http://www.rpi.edu/dept/arc/training/latex/resumes/)
% It was modified by Katy Huff
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%    PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{letter} % Default font size of the document, change to 10pt to fit more text
\usepackage{graphicx}
%\usepackage{newcent} % Default font is the New Century Schoolbook PostScript font
%\usepackage{helvet} % Uncomment this (while commenting the above line) to use the Helvetica font

% Margins
\usepackage[left=1.25in,right=1.25in,top=1in,bottom=1in]{geometry}
%\let\raggedleft\raggedright % Pushes the date (at the top) to the left, comment this line to have the date on the right

\usepackage{eso-pic,graphicx}

%--------------------------------------------------------------------------------------
%--------INPUT DATA
%--------------------------------------------------------------------------------------
\usepackage{xspace}
\newcommand{\StudentFirstName}{Andrei\xspace}
\newcommand{\StudentLastName}{Rykhlevskii\xspace}
\newcommand{\RecipientName}{Profs. Ghiaasiaan, Pazsit, and Pozzi\xspace}
\newcommand{\RecipientAddress}{Executive Editors\\Annals of Nuclear Energy}
%\newcommand{\<++>}{<++>\xspace}

\begin{document}
\AddToShipoutPictureBG*{\includegraphics[width=\paperwidth,height=\paperheight]{background.pdf}}

%----------------------------------------------------------------------------------------
%    ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{\RecipientName\\
        \RecipientAddress\xspace}

%-------------------------------------------------------------------------------
%    YOUR NAME & ADDRESS SECTION
%-------------------------------------------------------------------------------
\address{Kathryn D. Huff\\
kdhuff@illinois.edu\\
118 Talbot Laboratory\\
104 Wright Street\\
MC-234\\
Urbana, IL 61801}

%-------------------------------------------------------------------------------
%    LETTER CONTENT SECTION
%-------------------------------------------------------------------------------

\opening{To \RecipientName,}

Please find enclosed a manuscript entitled: ``Modeling and Simulation of Online
Reprocessing in the Thorium-Fuelled Molten Salt Breeder Reactor'' which Mr. 
Andrei Rykhlevskii and I are submitting for exclusive consideration of 
publication as a research article in Annals of Nuclear Energy.

This manuscript describes and demonstrates a computational modeling capability
for Molten Salt Reactor (MSR) online reprocessing analysis.  A Python package,
SaltProc, has been developed to simulate MSR online reprocessing by modeling
the changing isotopic composition of an irradiated fuel salt. The simulation
workflow additionally leverages the SERPENT 2 Monte Carlo software for neutron
transport and depletion calculations. An introduction to SaltProc capabilities
is followed by demonstration in the context of online reprocessing in the
Molten Salt Breeder Reactor (MSBR).

Thank you for your consideration of our work.  I expect it will be of interest
to a broad readership concerned with multi-physics simulation of next generation
reactors.  Please address all correspondence concerning this manuscript to me at the University of Illinois.

\closing{Sincere regards,
\includegraphics[height=2cm]{signature}\\
        \fromsig{Kathryn Huff\\
Assistant Professor\\
Nuclear, Plasma, \& Radiological Eng.\\
U. Illinois at Urbana-Champaign}
}

%-------------------------------------------------------------------------------

\end{letter}

\end{document}