### 2018-05-10 conclusion.tex

\section{Conclusion}

% The main conclusions of the study may be presented in a short Conclusions
% section, which may stand alone or form a subsection of a Discussion or Results
% and Discussion section.

\FloatBarrier