### 2018-05-18 conclusion.tex
Paper outline finished

\section{Conclusion}

Condensed copy-paste of the thesis conclusion.
\begin{itemize}
    \item Full-core model which is better the most exists. Importance of full-core approach for multi-region designs.
    \item K-eff dynamics and explanation of this dynamics.
    \item Spectral shift explanation.
    \item Why spectral shift causes safety parameters worsening and power profile changes.
    \item $^{232}$Th rate is in a good agreement with references.
\end{itemize}

\FloatBarrier