### 2018-05-31 discussion.tex

\section{Discussion}
% This should explore the significance of the results of the work, not repeat
% them. A combined Results and Discussion section is often appropriate. Avoid
% extensive citations and discussion of published literature.
Short 2-3 paragraph discussion. This will emphasize:
\begin{itemize}
  \item Discuss reason of spectral shift (heavy \gls{FP} accumulation, Pu isotopes).
  \item Neutron spectral shift which causes safety parameters worsening.  
  \item Couple words about power profile.
\end{itemize}