### 2018-08-31 abstract.tex
Ansh edits incorporated #21

\begin{abstract}

Today,  climate change is driving the search for new ways to generate 
carbon-free, reliable base-load power. Thus,
interest in advanced nuclear energy and particularly \glspl{MSR} has resurged, 
with multiple new companies pursuing commercialization of \gls{MSR} 
designs (e.g. Transatomic, Terrapower, Terrestrial, Moltex Energy, Thorcon). To further 
develop these \gls{MSR} concepts, researchers need simulation tools for analyzing liquid 
fueled \gls{MSR} depletion and fuel processing. However, most 
contemporary nuclear reactor physics software is unable to perform high-fidelity 
full-core depletion calculations for a reactor design with online reprocessing. This paper 
introduces a Python package, SaltProc, which couples with the Monte Carlo code, 
SERPENT2, to simulate \gls{MSR} online reprocessing by modeling the changing 
isotopic composition of \gls{MSR} fuel salt. SaltProc capabilities were 
demonstrated for full-core high-fidelity model of the commercial \gls{MSBR} 
concept and compared well to previous results from lower-fidelity analyses.

\end{abstract}